---
title: "Simple linear Regression and Bayesian Linear Regression"
author: "Kavinda Uthsuka"
date: "3/29/2022"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Importing the dataset
```{r}
library(BAS)
data(bodyfat)
summary(bodyfat)
```
We will construct a Bayesian model of simple linear regression, which uses Abdomen to predict response variable bodyfat.

Fit frequentist OLS linear regression
```{r}
bodyfat.lm = lm(Bodyfat ~ Abdomen,data = bodyfat)
summary(bodyfat.lm)
```
Regression line:
  Bodyfat = -39.28018 + 0.63130*Abdomen

Bodyfat will increase 0.6313 units when every additional increment of 1cm of Abdomen.


Extract coefficient
```{r}
beta = coef(bodyfat.lm)
beta
```

Visualize regression line on the scatter plot
```{r}
library(ggplot2)
ggplot(data = bodyfat,aes(x=Abdomen,y=Bodyfat))+
  geom_point(color = 'blue')+
  geom_abline(intercept = beta[1],slope=beta[2],size=1)+
  xlab("abdomen circumference (cm) ")
```

Calculate the Mean Squared Error (MSE)
```{r}
resid = residuals(bodyfat.lm)
n = length(resid)

MSE = 1/(n-2) * sum((resid ^2))
MSE
```
We apply the scatterplot of residuals versus fittedvalues, which provides an additonal isual check of the model adequacy.
```{r}
#Combine resiaudals and fitted values into a data frame
result = data.frame(fitted_values = fitted.values(bodyfat.lm), residuals = residuals(bodyfat.lm))

#load library and plot residuals versus fitted values
ggplot(data = result,aes(x= fitted_values,y=residuals))+
  geom_point(pch = 1,size = 2) + 
  geom_abline(intercept = 0,slope = 0)+
  xlab(expression(paste("fitted value",widehat(Bodyfat))))+
  ylab("residuals")
```
There is one outlier point. Therefore we have to deal with it.

Find the observation with the largest fitted value.
```{r}
which.max(as.vector(fitted.values(bodyfat.lm)))
```
Shows this observation has the largest Abdomen
```{r}
which.max(bodyfat$Abdomen)
```
Let's plot normal probability plot of the residuals for check the assumption of normally distributed errors
```{r}
plot(bodyfat.lm,which = 2)
```

Credible intervals for slope beta and y intercept alpha
```{r}
output = summary(bodyfat.lm)$coef[,1:2]
output
```
```{r}
out = cbind(output,confint(bodyfat.lm))
colnames(out) = c("posterior mean","posterior std","2.5","97.5")
round(out,2)
```
Let's plot the prediction intervals
```{r}
#construct the current prediction
alpha = bodyfat.lm$coefficients[1]
beta = bodyfat.lm$coefficients[2]
new_x = seq(min(bodyfat$Abdomen),max(bodyfat$Abdomen), length.out = 100)

y_hat = alpha + beta*new_x
#Get lower and upper bounds for mean
ymean = data.frame(predict(bodyfat.lm, newdata = data.frame(Abdomen = new_x), interval = "confidence",level = 0.95))

#Get lower and upper bounds for prediction
ypred = data.frame(predict(bodyfat.lm, newdata = data.frame(Abdomen = new_x), interval = "prediction",level = 0.95))

output = data.frame(x = new_x,y_hat = y_hat,ymean_lwr = ymean$lwr , ymean_upr = ymean$upr, ypred_lwr = ypred$lwr , ypred_upr = ypred$upr)

#Extract potential outlier data point
outlier = data.frame(x = bodyfat$Abdomen[39],y=bodyfat$Bodyfat[39])

#scatterplot of original
plot1 = ggplot(data = bodyfat, aes(x=Abdomen,y=Bodyfat)) + geom_point(color = "blue")

# Add bounds of mean and prediction
plot2 = plot1 + geom_line(data=output,aes(x=new_x,y=y_hat,color="first"),lty = 1)+
  geom_line(data=output,aes(x=new_x,y=ymean_lwr,lty = "second")) +
  geom_line(data=output,aes(x=new_x,y=ymean_upr,lty = "second")) +
  geom_line(data=output,aes(x=new_x,y=ypred_upr,lty = "third")) +
  geom_line(data=output,aes(x=new_x,y=ypred_lwr,lty = "third")) + 
  scale_colour_manual(values = c("orange"),labels = "Posterior mean",name = "") +
  scale_linetype_manual(values = c(2,3),labels = c("95% CI for mean","95% CI for predictions"),name = "")+
  theme_bw()+
  theme(legend.position = c(1,0),legend.justification = c(1,5,0))
#Identify potential outlier
plot2 + geom_point(data = outlier , aes(x=x,y=y),color="orange",pch = 1,cex=6)

```


```{r}
pred.39 = predict(bodyfat.lm,newdata = bodyfat[39,], interval = "prediction", level = 0.95)
out = cbind(bodyfat[39,]$Abdomen,pred.39)
colnames(out) = c("abdomen","prediction","lower","upper")
out
```
The bayesian posterior distribution results of alpha and beta show that under the reference prior, the posterior credible intervals are numerically equivalent to the cionfidence intervals from the classical frequentist OLS analysis.

Since the credible intervals are numerically the same as the confidence intervals, we use the lm function to obtain the OLS estimates and construct the  credible intervals of alpha and beta.

```{r}
output1 = summary(bodyfat.lm)$coef[,1:2]
output1

out1 =cbind(output1,confint(bodyfat.lm))
colnames(out1) = c("posterior mean","posterior std","2.5","97.5")
round(out1,2)
```
We believe that there is 95% chance that body fat will increase by 58% up to 69% for every additional 10 centimeter increase in the waist circumference.

